Bayesian multi-proxy reconstruction of early Eocene latitudinal temperature gradients

### 2023-09-22 manuscript/main.qmd
model validation method

We created test data from these gradients as follows: To capture both limited sample size and skewed geographic origins of samples, we randomly sampled (200 iterations) the latitudes of the EECO data set (n = 34) used in this study, at sample sizes of 5, 10 and 20, and also used the full set of 34 latitudes. For each latitude, we inferred a location mean temperature from the gradients, adding random noise from a normal distribution with a standard deviation of 2.2, which is the residual standard deviation of modern sea surface temperature averages from the modelled modern gradient (see below). We then randomly allocated a sample size $s$ from the sample sizes present in the EECO data set to each location, and randomly generated $s$ temperature data from a normal distribution with mean equal to the previously generated location mean temperatures, and a standard deviation of 3.8. This standard deviation is the average standard deviation of the EECO geochemical proxy data, and with that we aim to simulate randomly distributed errors in the proxy data. This approach should give a fair assessment of how the model setup can be affected by the known unknowns (quantifiable errors), but we acknowledge that it cannot quantify the potential impact of unknown, systematic offsets that may bias all proxy data from one location in the same direction.

>>>>

We created test data from these gradients as follows: We randomly sampled (1000 iterations) latitudes at sample sizes of 5, 10 and 20, with the probability of a latitude being sampled scaling with the decreasing surface area towards higher latitudes; i.e. lower latitudes are sampled more frequently. For the largest sample size (n = 34), we used the latitudes of the EECO data set of this study in all iterations. For each latitude, we inferred a location mean temperature from the gradients, adding random noise from a normal distribution with a standard deviation of 3.8, which is the average standard deviation of the EECO geochemical proxy data. With that, we aim to simulate randomly distributed errors in the proxy data, which could arise from miscalibrations, measurement errors, seasonal effects, ect. We acknowledge that this approach cannot quantify the potential impact of systematic offsets that may bias all proxy data in the same direction, nor do we know whether a standard deviation of 3.8 is actual, average magnitude of uncertainty that the proxy compilation is afflicted with.
